### 2017-04-11 CS798-ConvexityAndOptimization/assignments/project.tex
proof for first alg

\end{proof}
\section{Ad Auctions}
\subsection{Single Slot}

>>>>

Part $1.$ is quite clear. Whenever a primal constraint is introduced, the While loop of Step $2$ does not terminate until the constraint is satisfied. Subsequent iterations do not decrease the variables of $x$, so the constraint remains satisfied.
To demonstrate $2.$ we will study a single iteration of the While loop. In an iteration of Step $2$ the change in $\sum_{j=1}^m y_j$ is precisely $1$. The change in the primal objective value $c^Tx$ is (from the update in Step $2a$):
$$\sum_{i \in S(j)} c_i (\frac{x_i}{c_i} + \frac{1}{|S(j)| c_i}) = \sum_{i \in S(j)}(x_i + \frac{1}{|S(j)|}) \leq 2$$
since $\sum_{i \in S(j)} x_i \leq 1$ at the time Step $2a$. Therefore $2.$ holds.
Let $i \in [n]$. We consider the dual constraint corresponding to $i$. Whenever we increase a dual variable $y_j$ such that $i \in S(j)$ by $1$ in Step $2b$ we had just increased $x_i$ in Step $2a$. We claim that 
$$x_i \geq \frac{1}{d}((1+\frac{1}{c_i})^{\sum_{j : i \in S(j)} y_j} - 1).$$
We will proceed by induction on the iterations of $\cA$. In the base case $x_i = 0$, and $\sum_{j : i \in S(j)} y_j = 0$ so the claim holds. Now consider an iteration in which some dual variable $y_k$ with $k \in S(j)$ is increased. Let $x_i'$ denote the value of $x_i$ after this iteration (and $x_i$ denote the value before).
x'_i &= x_i(1 + \frac{1}{c_i}) + \frac{1}{|S(j)|c_i} \\
&\geq x_i(1+\frac{1}{c_i}) + \frac{1}{d\cdot c_i} \\
&\geq \frac{1}{d}((1+\frac{1}{c_i})^{\sum_{j : i \in S(j)\backslash\{k\}}y_j} - 1)(1 + \frac{1}{c_i})^y_k + \frac{1}{d\cdot c_i} &\text{by the inductive hypothesis on $x_i$} \\
&= \frac{1}{d}((1+\frac{1}{c_i})^{\sum_{j :i \in S(j)} y_j} -1).
So the claim holds by induction. Now we compute an upper bound on $x'_i$. By the condition of the While loop, $x_i <1$. We also have $c_i \geq 1$ and $d\geq 1$, so
$$x'_i =  x_i(1 + \frac{1}{c_i}) + \frac{1}{|S(j)|c_i} \leq 1(1+1) + 1 = 3.$$
Combining inequalities we have
$$ \frac{1}{d}((1+\frac{1}{c_i})^{\sum_{j :i \in S(j)} y_j} -1) \leq 3.$$
Using once more that $c_i \geq 1$ and rearranging for $\sum_{j : in \in S(j)} y_j$ we obtain
$$\sum_{j : i \in S(j)} y_j \leq c_i \log(3d+1) = c_i \cdot O(\log d)$$
as desired.
It remains to use these claims to verify the competitive factors. This is done via duality of linear programs. Let $p^*$ be the optimal value of $(P)$. Let $f^*$ be the optimal value of $(D)$. By claim $2$ we have
$$c^Tx \leq 2 \sum_{j} y_j.$$
Since $x$ is primal feasible, $c^Tx \geq p^*$. So 
$$2\sum_j y_j \geq c^Tx \geq p^* = f^*.$$
Thus $y$ is $2$-competitive. By claim $3$, $y' \in \R^m$ where $y'_j = y_j/O(\log d)$ is feasible for $(D)$.
So similarly,
$$c^Tx \leq 2 O(\log d) \sum_j y_j \leq O(\log d) f^* = O(\log d) p^*.$$
Therefore $x$ is $O(\log d)$-competitive.
\end{proof}
\section{Ad Auctions}
\subsection{Single Slot}
